---
title: "Prediction Model of Quantified Self Movement"
author: "Thammanoon Kawinfruangfukul"
output: html_document
---
## Synopsis
This report describes how a model is built for predicting the manner in which six participants performed barbell lifts. The data were gathered from accelerometers on the belt, forearm, arm, and dumbell [1]. Exploratory data analysis, model building, and model evaluation are described below.

## Exploratory Data Analysis
This section performs cleaning data and extracting important information for building a prediction model.
```{r, echo=TRUE, cache=TRUE, warning=FALSE}
# Loading required libraries
library(caret)
library(parallel)
library(doParallel)
set.seed(1000)

# Loading and preprocessing training and testing data sets
cpath        <- paste(getwd(), "/data/pml-training.csv", sep = "");
pml_training <- read.csv(cpath, header = TRUE, sep = ",", na.strings = c("NA","#DIV/0!",""))

cpath        <- paste(getwd(), "/data/pml-testing.csv", sep = "");
pml_testing  <- read.csv(cpath, header = TRUE, sep = ",", na.strings = c("NA","#DIV/0!",""))

# Preprocessing: Subsetting the data sets (excluding user and time-related information)
training <- pml_training[, -c(1:7)]
testing  <- pml_testing[, -c(1:7)]

# Preprocessing: Removing a column with NA
training  <- training[, colSums(is.na(training)) == 0]
testing   <- testing[, colSums(is.na(testing)) == 0]
```

## Model Building
The data are divided into training and test sets (80%:20%). The prediction model is random forest. In the model building, principal components analysis (PCA) is used for grouping quantitative variables that are highly collarated with each other. Moreover, cross validation (10-fold) is applied for picking variables to include in the model. Both PCA and cross validation are configured in fitControl variable.

Random forest is used because it provides a high-accuracy model; however, it may introduce overfitting and has a high out of sample error. PCA and cross validation help in alleviating the overfitting and reducing the out of sample error.
```{r, echo=TRUE, cache=TRUE, warning=FALSE}
inTrain      <- createDataPartition(training$classe, p = 0.80, list = F)
training_set <- training[inTrain,]
test_set     <- training[-inTrain,]

# Configuring parallel processing
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)

# Configuring trainControl object
fitControl <- trainControl(method = "cv", number = 10, preProcOptions="pca", allowParallel = TRUE)

# Developing training model
fit <- train(classe ~ ., method="rf", data = training_set, trControl = fitControl)

# De-registering parallel processing cluster
stopCluster(cluster)

# Show the model
fit
confusionMatrix.train(fit)
```

## Model Evaluation
```{r, echo=TRUE, cache=TRUE, warning=FALSE}
# Predict test set
pred <- predict(fit, test_set)

# Show confusion matrix and statistics
accuracy <- confusionMatrix(pred, test_set$classe)$overall['Accuracy']
confusionMatrix(pred, test_set$classe)
```
Using the prediction model to evaluate test set, we find that the model has `r accuracy` accuracy. 

We will continue evaluating out of sample data by using the 20 samples given in pml-testing.csv. The following are prediction results. The prediction results match with the results from Course Project Prediction Quiz.
```{r, echo=TRUE, cache=TRUE, warning=FALSE}
predict(fit, testing)
```

## Reference
[1] Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
